Multi-level modeling with Openflexo/\FML - A contribution to the \mpc

### 2021-05-31 6-modelpresentation.tex
sync

Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and are executable by the \FML execution engine.

@@

We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections. A first tool offers a graphical edition of a Process type. The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.

@@

The other tool, called \textit{Enacted process graphical editor} and shown on figure~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed. This tool is represented on the right side of the figure~\ref{fig:ToolingArchitecture}. It follows the same architecture pattern of the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram. Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process. All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith"). All tasks get a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.

\begin{figure*}
 \centering
     \includegraphics[width=1.0 \textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
     \caption{Screenshot of the enacted process graphical editor}
    \label{fig:ScreenshotEnactedProcessEditor}
\end{figure*}

>>>>

Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and can be executed by the \FML execution engine.

@@

We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections. A first tool offers a graphical edition of a Process type. The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.

@@

The other tool, called \textit{Enacted process graphical editor} and shown on figure~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed. This tool is represented on the right side of the figure~\ref{fig:ToolingArchitecture}. It follows the same architecture pattern of the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram. Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process. All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, related task is automatically assigned to \inst{Ann Smith}). Each task has a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.

\begin{figure*}
 \centering
     \includegraphics[width=\textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
     \caption{Screenshot of the enacted process graphical editor}
    \label{fig:ScreenshotEnactedProcessEditor}
\end{figure*}

### 2021-05-31 7-requirementssatisfaction.tex
sync

In the previous section~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements. In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications. We classify those techniques into three categories: 
\begin{itemize}
    \item \textbf{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
    \item\textbf{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
    \item \textbf{tooling}: the requirements are enforced by the implementation of the associated tooling.
\end{itemize}

@@

\begin{itemize}
    \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
    \item \textbf{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
    \item \textbf{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
    \item \textbf{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
    \item \textbf{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
\end{itemize}

>>>>

In the previous section~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements. In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications. We classify these techniques into three categories: 
\begin{itemize}
    \item \emph{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
    \item \emph{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
    \item \emph{tooling}: the requirements are enforced by the implementation of the associated tooling.
\end{itemize}

@@

\begin{itemize}
    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
    \item \emph{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
    \item \emph{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
    \item \emph{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
\end{itemize}
